---
title: "Visualizations & Analysis"
author: "George D. Quinn"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
## Hand coded to Major Catagories

```{r setup, include=FALSE}
library(readxl)
library(dplyr)

# Step 1: Read the Excel file
hand_database <- read_excel("Full_Database.xlsx", sheet = "hand_coded")

hand_database$mc1 <- apply(hand_database[, c("6", "7", "33", "4")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc2 <- apply(hand_database[, c("5", "10", "13")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc3 <- apply(hand_database[, c("11", "12", "14", "15", "16","17","18","19")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc4 <- apply(hand_database[, c("30", "31")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc5 <- apply(hand_database[, c("2", "3")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc6 <- apply(hand_database[, c("20", "21", "22", "24")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc7 <- apply(hand_database[, c("23")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc8 <- apply(hand_database[, c("26", "32")], 1, function(x) as.integer(any(x == 1)))

hand_database$mc9 <- apply(hand_database[, c("1", "8", "9", "27", "28","29","34","35")], 1, function(x) as.integer(any(x == 1)))
```

# STM to Major Catagories

```{r}
STM_data <- read_excel("Full_Database.xlsx", sheet = "STM_data")

STM_data$mc1 <- apply(STM_data[, c("6", "26", "28", "32")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc2 <- apply(STM_data[, c("4", "15", "22", "24", "33")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc3 <- apply(STM_data[, c("2", "9", "11", "17", "20","31")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc4 <- apply(STM_data[, c("13", "21", "25")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc5 <- apply(STM_data[, c("12", "27", "34")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc6 <- apply(STM_data[, c("1", "5", "18", "26", "30")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc7 <- apply(STM_data[, c("14","23","35")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc8 <- apply(STM_data[, c("3", "29")], 1, function(x) as.integer(any(x == 1)))

STM_data$mc9 <- apply(STM_data[, c("8", "16", "19")], 1, function(x) as.integer(any(x == 1)))
```

# Major Catagories STM - Hand codes correlation martix 
```{r}
library(ggplot2)
library(reshape2)
library(dplyr)

# Assuming 'hand_database' and 'STM_data' are already loaded and have mc1 to mc9 columns
# Subset the relevant columns (mc1 to mc9) from both datasets
hand_mc <- select(hand_database, starts_with("mc"))
STM_mc <- select(STM_data, starts_with("mc"))

# Bind the two sets of columns by row and rename them
combined_mc <- bind_rows(hand_mc, STM_mc) %>%
  rename(
    Experience = mc1,
    `Leadership Qualities` = mc2,
    `Personal Qualities` = mc3,
    `Government Management` = mc4,
    `Government Activity / Philosophy` = mc5,
    `Domestic Policies` = mc6,
    `Foreign Policies` = mc7,
    `Group Connections` = mc8,
    Miscellaneous = mc9
  )

# Calculate the correlation matrix
cor_matrix <- cor(combined_mc, use = "complete.obs", method = "pearson")

# Melt the correlation matrix into long format
melted_cor_matrix <- melt(cor_matrix)

# Plot the correlation matrix using ggplot2
cor_plot <- ggplot(data = melted_cor_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", value)), vjust = 1) +
  scale_fill_gradient2(low = "Purple", high = "green", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", name="Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(size = 12, margin = margin(t = 20)),
        axis.title.y = element_text(size = 12, margin = margin(r = 20)),
        axis.text.y = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "hand_coded", y = "stm_coded") +
  coord_fixed()

# Print the plot
print(cor_plot)
```

# All 35 Correlation Matrix

```{r}
# Merge the two data frames
unified_set <- merge(hand_database, STM_data, by = "Responses")

# View the first few rows of the unified data frame to verify it's correct
head(unified_set)

```

```{r}
unified_set_renamed <- unified_set %>%
  rename_with(~sub("^(\\d+)\\.y$", "STM_\\1", .), matches("\\.y$")) %>%
  rename_with(~sub("^(\\d+)\\.x$", "HAND_\\1", .), matches("\\.x$"))

# View the names of the updated data frame to verify
names(unified_set_renamed)

unified_set_renamed <- unified_set_renamed %>%
  rename_with(~sub("^mc(\\d+)\\.y$", "STM_mc\\1", .), matches("^mc\\d+\\.y$")) %>%
  rename_with(~sub("^mc(\\d+)\\.x$", "HAND_mc\\1", .), matches("^mc\\d+\\.x$"))

names(unified_set_renamed)
```

```{r}
# Assuming unified_set_renamed is already loaded and has columns STM_1 to STM_35 and HAND_1 to HAND_35

# Specify the HAND columns to include based on the file you provided
hand_columns <- c(21, 14, 32, 11, 25, 6, 8, 27, 11, 30, 15, 3, 33, 23, 10, 1, 11, 25, 20, 11, 13, 30, 23, 15, 13, 33, 2, 7, 3, 22, 12, 5, 13, 9, 23)

# Create a vector with the STM column names
stm_columns <- paste0("STM_", 1:35)

# Create a vector with the HAND column names using the hand_columns vector
hand_column_names <- paste0("HAND_", hand_columns)

# Select only the relevant columns from the unified_set_renamed data frame
stm_data <- unified_set_renamed[stm_columns]
hand_data <- unified_set_renamed[hand_column_names]

# Calculate the correlation matrix between the STM and HAND columns
correlation_matrix <- cor(stm_data, hand_data, use = "pairwise.complete.obs")

# Convert the correlation matrix to a long format for easy plotting or viewing
correlation_long <- reshape2::melt(correlation_matrix)

# View the correlation_long data frame
print(correlation_long)
```


```{r}
library(ggplot2)

# Assuming 'correlation_long' is already created with 'reshape2::melt()' from the correlation matrix

# Convert factor variables to character so that they are plotted in the original order
correlation_long$Var1 <- as.character(correlation_long$Var1)
correlation_long$Var2 <- as.character(correlation_long$Var2)

# Plot the correlation matrix
ggplot(correlation_long, aes(Var2, Var1, fill = value)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", value)), vjust = 1) +
  scale_fill_gradient2(low = "Purple", high = "green", mid = "white", midpoint = 0, limit = c(-0.5, 0.5), space = "Lab", name="Correlation") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.text.y = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Correlation", 
       title = "Correlation Matrix", 
       x = "HAND Columns", 
       y = "STM Columns")

# If Var1 and Var2 are not in the order you want, consider reordering them before plotting
```


# Data Set Totals 

```{r}
# Calculate frequencies for each of the columns mc1 to mc9
freq_mc1 <- table(hand_database$mc1, useNA = "ifany")
freq_mc2 <- table(hand_database$mc2, useNA = "ifany")
freq_mc3 <- table(hand_database$mc3, useNA = "ifany")
freq_mc4 <- table(hand_database$mc4, useNA = "ifany")
freq_mc5 <- table(hand_database$mc5, useNA = "ifany")
freq_mc6 <- table(hand_database$mc6, useNA = "ifany")
freq_mc7 <- table(hand_database$mc7, useNA = "ifany")
freq_mc8 <- table(hand_database$mc8, useNA = "ifany")
freq_mc9 <- table(hand_database$mc9, useNA = "ifany")

# Print the frequency tables
print(freq_mc1)
print(freq_mc2)
print(freq_mc3)
print(freq_mc4)
print(freq_mc5)
print(freq_mc6)
print(freq_mc7)
print(freq_mc8)
print(freq_mc9)
```

```{r}
# Calculate frequencies for each of the columns mc1 to mc9
STM_mc1 <- table(STM_data$mc1, useNA = "ifany")
STM_mc2 <- table(STM_data$mc2, useNA = "ifany")
STM_mc3 <- table(STM_data$mc3, useNA = "ifany")
STM_mc4 <- table(STM_data$mc4, useNA = "ifany")
STM_mc5 <- table(STM_data$mc5, useNA = "ifany")
STM_mc6 <- table(STM_data$mc6, useNA = "ifany")
STM_mc7 <- table(STM_data$mc7, useNA = "ifany")
STM_mc8 <- table(STM_data$mc8, useNA = "ifany")
STM_mc9 <- table(STM_data$mc9, useNA = "ifany")

# Print the frequency tables
print(STM_mc1)
print(STM_mc2)
print(STM_mc3)
print(STM_mc4)
print(STM_mc5)
print(STM_mc6)
print(STM_mc7)
print(STM_mc8)
print(STM_mc9)
```

```{r}
library(dplyr)
library(tidyr)

# Sample data frame creation for demonstration purposes
# Assuming unified_set_renamed has columns named HAND_mc1, HAND_mc2, ..., HAND_mc9, STM_mc1, STM_mc2, ..., STM_mc9
# unified_set_renamed <- data.frame(...)

# Initialize an empty list to store crosstab results
crosstabs_list <- list()

for(i in 1:9) {
  # Dynamically create column names
  hand_col <- paste0("HAND_mc", i)
  stm_col <- paste0("STM_mc", i)
  
  # Generate the crosstab for the current pair of columns
  crosstab_result <- table(unified_set_renamed[[hand_col]], unified_set_renamed[[stm_col]])
  
  # Store the result in the list with a descriptive name
  crosstabs_list[[paste0("Crosstab_", hand_col, "_", stm_col)]] <- crosstab_result
}

# Now crosstabs_list contains all the crosstabs
# You can access each crosstab using crosstabs_list[["Crosstab_HAND_mc1_STM_mc1"]] for the first pair, and so on.

# If you want to print all crosstabs
for (name in names(crosstabs_list)) {
  cat("Crosstab for:", name, "\n")
  print(crosstabs_list[[name]])
  cat("\n")
}
```

# STM - Hand Desciptives

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

# Assuming unified_set_renamed is your dataframe

# Reshape the data into a long format
long_format <- unified_set_renamed %>%
  pivot_longer(cols = starts_with("HAND_mc") | starts_with("STM_mc"),
               names_to = c("Type", "mc_number"),
               names_sep = "_",
               values_to = "value") %>%
  mutate(Type = factor(Type, levels = c("HAND", "STM")))

# Count the occurrences of 1s
counts <- long_format %>%
  filter(value == 1) %>%
  group_by(Type, mc_number) %>%
  summarise(count = n(), .groups = 'drop')

# Generate the side-by-side bar chart for counts of 1s
ggplot(counts, aes(x = mc_number, y = count, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Major Category - Hand codes & STM",
       x = "Major Category",
       y = "Coded Responses") +
  scale_fill_manual(values = c("HAND" = "skyblue", "STM" = "tomato")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## HAND - STM MAJOR CAT - Correlations 
```{r}
library(ggplot2)
library(dplyr)

# Assuming 'counts' is your dataframe that contains all the data
# Filter out HAND counts
hand_counts <- counts %>% 
  filter(Type == "HAND") %>% 
  select(mc_number, count)

# Filter out STM counts
stm_counts <- counts %>% 
  filter(Type == "STM") %>% 
  select(mc_number, count)

# Assuming you need to correlate the counts by 'mc_number' now:
# First, ensure both dataframes are ordered by 'mc_number' to align them properly
hand_counts <- hand_counts %>% arrange(mc_number)
stm_counts <- stm_counts %>% arrange(mc_number)

# Combine the counts into one dataframe by 'mc_number'
# The mc_number column should be identical between the two sets, so we can bind the rows
combined_counts <- full_join(hand_counts, stm_counts, by = "mc_number") %>%
  rename(hand_count = count.x, stm_count = count.y)

# Calculate the correlation on the combined data
cor_value <- cor(combined_counts$hand_count, combined_counts$stm_count, use = "complete.obs")

# Create a scatter plot with a regression line to visualize the correlation
cor_plot <- ggplot(combined_counts, aes(x = hand_count, y = stm_count)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = paste("Correlation between HAND and STM counts (r =", round(cor_value, 2), ")"),
       x = "HAND Count",
       y = "STM Count") +
  theme_minimal()

# Print the plot
print(cor_plot)
```

## STM - HAND MAJOR CAT: Correlations

```{r}
# Specify the HAND columns to include based on the file you provided
hand_mc_columns <- paste0("HAND_mc",1:9)

# Create a vector with the STM column names
stm_mc_columns <- paste0("STM_mc", 1:9)

# Create a vector with the HAND column names using the hand_columns vector
hand_column_names <- paste0("HAND_", hand_columns)

# Select only the relevant columns from the unified_set_renamed data frame
stm_data_mc <- unified_set_renamed[stm_mc_columns]
hand_data_mc <- unified_set_renamed[hand_mc_columns]

# Calculate the correlation matrix between the STM and HAND columns
correlation_matrix_mc <- cor(stm_data_mc, hand_data_mc, use = "pairwise.complete.obs")

# Convert the correlation matrix to a long format for easy plotting or viewing
correlation_major_cat <- reshape2::melt(correlation_matrix_mc)

# Check the renamed data frame
print(correlation_major_cat)
```

# STM - HAND : tile chart

```{r}
library(ggplot2)
library(dplyr)

# Assuming correlation_major_cat is the original data frame

# Define the new names for the HAND and STM variables
hand_names <- c("Experience", "Leadership Qualities", "Personal Qualities", "Gov't Management", 
                "Gov't Philosophy", "Domestic Policies", "Foreign Policies", 
                "Group Connections", "Miscellaneous")
stm_names <- paste(hand_names, "STM", sep=" ")

# Use scale_x_discrete and scale_y_discrete to rename the axis ticks in the plot
tile_chart <- ggplot(correlation_major_cat, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", value)), vjust = 1) +
  scale_fill_gradient2(low = "purple", high = "green", mid = "white", midpoint = 0, limit = c(-0.5, 0.5), space = "Lab", name="Correlation") +
  scale_x_discrete(labels = hand_names) +
  scale_y_discrete(labels = stm_names) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(size = 12, margin = margin(t = 20)),
        axis.title.y = element_text(size = 12, margin = margin(r = 20)),
        axis.text.y = element_text(angle = 45, hjust = 1),) +
  labs(fill = "Correlation", 
       title = "STM-HAND Correlation Matrix", 
       x = "HAND Columns", 
       y = "STM Columns") +
  coord_fixed()
# Display the tile chart
print(tile_chart)
```


```{r}

```



```{r}
library(dplyr)
library(stargazer)

# Assuming unified_set_renamed is your dataframe

# Reshape the data into a long format
long_format <- unified_set_renamed %>%
  pivot_longer(cols = starts_with("HAND_mc") | starts_with("STM_mc"),
               names_to = c("Model", "Category"),
               names_sep = "_",
               values_to = "Value") %>%
  mutate(Model = factor(Model, levels = c("HAND", "STM")))

# Summarize the counts of 1s for each category and model
frequency_table <- long_format %>%
  filter(Value == 1) %>%
  count(Model, Category) %>%
  spread(key = Category, value = n, fill = 0) # Spread categories across columns

print(frequency_table)
```

# STM - HAND confusion Matrix

```{r}
# Plot the heatmap
ggplot(correlation_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", name="Correlation") +
  theme_minimal() +
  labs(title = "Correlation between HAND_mcX and STM_mcX", x = "MC Category", y = "") +
  geom_text(aes(label = ifelse(is.na(value), "NA", sprintf("%.2f", value))), color = "black", size = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Importing Pilot GPT-4 derived data 

```{r}
library(readxl)
AI_unified <- read_excel("AI_unified.xlsx")
```

# HAND - GPT-4 Major Categories 

```{r}
# Specify the HAND columns to include based on the file you provided
hand_mc_columns <- paste0("HAND_mc",1:9)

# Create a vector with the STM column names
AI_mc_columns <- paste0("AI_mc", 1:9)

# Create a vector with the HAND column names using the hand_columns vector
hand_column_names <- paste0("HAND_", hand_columns)

# Select only the relevant columns from the unified_set_renamed data frame
AI_data_mc <- AI_unified[AI_mc_columns]
hand_data_mc <- AI_unified[hand_mc_columns]

# Calculate the correlation matrix between the STM and HAND columns
correlation_matrix_mc_AI <- cor(hand_data_mc, AI_data_mc, use = "pairwise.complete.obs")

# Convert the correlation matrix to a long format for easy plotting or viewing
correlation_major_cat_AI <- reshape2::melt(correlation_matrix_mc_AI)

# Check the renamed data frame
print(correlation_major_cat_AI)
```

# GPT4-HAND Major Cat Visualizations

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

# Assuming unified_set_renamed is your dataframe

# Reshape the data into a long format
long_format_AI <- AI_unified %>%
  pivot_longer(cols = starts_with("HAND_mc") | starts_with("AI_mc"),
               names_to = c("Type", "mc_number"),
               names_sep = "_",
               values_to = "value") %>%
  mutate(Type = factor(Type, levels = c("HAND", "AI")))

# Count the occurrences of 1s
counts_AI <- long_format_AI %>%
  filter(value == 1) %>%
  group_by(Type, mc_number) %>%
  summarise(count = n(), .groups = 'drop')

# Generate the side-by-side bar chart for counts of 1s
ggplot(counts_AI, aes(x = mc_number, y = count, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Major Category - Hand codes & AI",
       x = "Major Category",
       y = "Coded Responses") +
  scale_fill_manual(values = c("HAND" = "skyblue", "AI" = "tomato")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
library(ggplot2)
library(dplyr)

# Assuming correlation_major_cat is the original data frame

# Define the new names for the HAND and STM variables
hand_names <- c("Experience", "Leadership Qualities", "Personal Qualities", "Gov't Management", 
                "Gov't Philosophy", "Domestic Policies", "Foreign Policies", 
                "Group Connections", "Miscellaneous")
AI_names <- paste(hand_names, "AI", sep=" ")

# Use scale_x_discrete and scale_y_discrete to rename the axis ticks in the plot
tile_chart_AI <- ggplot(correlation_major_cat_AI, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", value)), vjust = 1) +
  scale_fill_gradient2(low = "purple", high = "green", mid = "white", midpoint = 0, limit = c(-0.5, 0.5), space = "Lab", name="Correlation") +
  scale_x_discrete(labels = hand_names) +
  scale_y_discrete(labels = AI_names) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(size = 12, margin = margin(t = 20)),
        axis.title.y = element_text(size = 12, margin = margin(r = 20)),
        axis.text.y = element_text(angle = 45, hjust = 1),) +
  labs(fill = "Correlation", 
       title = "AI-HAND Correlation Matrix", 
       x = "HAND Columns", 
       y = "AI Columns") +
  coord_fixed()

# Display the tile chart
print(tile_chart_AI)
```



# HAND - GPT-4 ALL 35
```{r}
# Assuming unified_set_renamed is already loaded and has columns STM_1 to STM_35 and HAND_1 to HAND_35

# Specify the HAND columns to include based on the file you provided
hand_columns <- paste0("HAND_", 1:35)

# Create a vector with the STM column names
AI_columns <- paste0("AI_", 1:35)

# Create a vector with the HAND column names using the hand_columns vector
hand_column_names <- paste0("HAND_", hand_columns)

# Select only the relevant columns from the unified_set_renamed data frame
AI_data <- AI_unified[AI_columns]
hand_data <- AI_unified[hand_columns]

# Calculate the correlation matrix between the STM and HAND columns
correlation_matrix_AI <- cor(AI_data, hand_data, use = "pairwise.complete.obs")

# Convert the correlation matrix to a long format for easy plotting or viewing
correlation_AI <- reshape2::melt(correlation_matrix_AI)

# View the correlation_long data frame
print(correlation_AI)
```


```{r}
library(ggplot2)

# Assuming 'correlation_long' is already created with 'reshape2::melt()' from the correlation matrix

# Convert factor variables to character so that they are plotted in the original order
correlation_AI$Var1 <- as.character(correlation_AI$Var1)
correlation_AI$Var2 <- as.character(correlation_AI$Var2)

# Plot the correlation matrix
ggplot(correlation_AI, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", value)), vjust = 1) +
  scale_fill_gradient2(low = "Purple", high = "green", mid = "white", midpoint = 0, limit = c(-0.5,0.5), space = "Lab", name="Correlation") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.text.y = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Correlation", 
       title = "Correlation Matrix", 
       x = "HAND Columns", 
       y = "STM Columns")
```

